%#!latex main

\chapter*{Acknowledgment}
\addcontentsline{toc}{chapter}{Acknowledgment}
\markboth{Acknowledgment}{Acknowledgment}

First I would like to thank deeply my supervisor, Professor Yahiko
Kambayashi at Kyoto University.  He gave me the opportunity to study
databases and workflow management systems, and continuously suggested
many helpful ideas since then.  He also extensively reviewed this thesis
and suggested many improvements.

I would also express my thanks to Professor Katsumasa Watanabe at Nara
Institute of Science and Technology (NAIST)\@.  He continuously supported
my study in Kyoto University and gave me the best environment for my
doctoral research.

I would like to thank Professor Shuzo Yajima at Kyoto University.  I had
studied with him at master course of Kyoto University.  He also
introduced me to Professor Kambayashi for carrying out my wishes to
begin researches about databases.

Associate Professor Kazumasa Yokota at Kyoto University discussed about
almost all part of this thesis and gave many helpful perspectives to me
about formalization of workflow model.  Workflow model in Chapter
\ref{chap:wfbase} would not be constructed without the discussions with
him.  I would like to acknowledge his contribution.

Associate Professor Shinji Kimura, Mr.\ Kazuyoshi Takagi, and the members
of Watanabe Laboratory at NAIST, cooperated to give me a good
environment for my study.

There are many other people I would like to thank.  I discussed many
ideas with Associate Professor Hiroshi Nunokawa at Miyagi University of
Education and the members of his research group, especially Mr.\ Masahiro
Hiji and Mr.\ Norihisa Segawa.  They have the similar research directions
with me in the area of computer-supported cooperative work, information
media, and databases.  Mr.\ Segawa now studies with me at NAIST\@.
Associate Professor Tsuneo Ajisaka at Kyoto University, Dr.\ Hiroyuki
Tarumi at NEC Corp., and Mr.\ Isao Kojima at Electrotechnical Laboratory
(ETL) suggested many useful comments from the viewpoints of software
engineering, groupware, and active database systems, respectively.
Dr.\ Hiroki Takakura at NAIST has helped me both in Kambayashi Laboratory
and in NAIST\@.  Associate Professor Masatoshi Arikawa (now at Hiroshima
City University), Ms.\ Masako Watanabe (Kyoto University), Mr.\ Shintaro
Meki (now at Okayama Prefectural University), Mr.\ Hideyuki Takada (now
at Mitsubishi Electric Corp.)\ also helped me when I was in Kambayashi
Laboratory.  The members of Kambayashi Laboratory offered the computer
environments to me every time I go to Kyoto University.

Finally, I would like to thank all people to help me, especially my
family and my friends in Choir Hamoru KOBE.

