%#!latex main

\chapter*{List of Publications by the Author}
\addcontentsline{toc}{chapter}{List of Publications by the Author}
\markboth{List of Publications by the Author}{List of Publications by
the Author}

\section*{Books}
\begin{enumerate}
 \item Y. Kambayashi, M. Arikawa, T. Kunishima, S. Konomi, H. Takada,
       and Y. Miyauchi.
       Distributed and Cooperative Media Series 4 --- Hypermedia and 
       Objectbase.
       Published from Kyoritsu Pub.\ Co.\ Ltd.
       Nov.\ 1995 (in Japanese).
\end{enumerate}

\section*{Major Publications}
\begin{enumerate}
 \item T. Naruse, T. Saito, and T. Kunishima.
       On an Algorithm Obtaining an Approximated Minimum Ball Containing
       $n$ Balls.
       Trans.\ IEICE, D-II, Vol.\ J73-D-II, No.\ 10, pp.\ 1792-1795,
       Oct.\ 1990
       (in Japanese).
 \item Y. Kambayashi, Q. Chen, and T. Kunishima.
       Coordination Manager: A Mechanism to Support Cooperative Work on
       Database Systems.
       Proc.\ 2nd Far-East Workshop on Future Database Systems,
       pp.\ 176-183,
       Apr.\ 1992.
 \item T. Kunishima and K. Yokota.
       Flexible Workflow Framework for Supporting Collaborative Works.
       Proc.\ International Symposium on Cooperative Database Systems for 
       Advanced Applications.
       pp.\ 501--508,
       Dec.\ 1996.
 \item K. Yokota, T. Kunishima, and H. Nakanishi.
       A Framework of Query Processing for Workflow Management.
       Lecture Notes in Artificial Intelligence,
       1997 (to appear).
 \item T. Kunishima and K. Yokota.  Workflow Base: A Flexible Workflow
       Model with Database Technologies.  Under submission in
       Trans. IPSJ (in Japanese).
\end{enumerate}

\section*{Technical Reports}
\begin{enumerate}
 \item T. Kunishima, S. Matsumoto, H. Ogino, H. Hiraishi, and
       S. Yajima.
       Computer-Aided Design using Multi Screen Graphic
       {MCMS} System.
       Proc.\ IPSJ Graphics and CAD Symposium, pp.\ 227-236,
       Nov.\ 1989 (in Japanese).
 \item T. Naruse, T. Kunishima, and T. Saito.
       Consideration on Setting up Bounding Volumes.
       IEICE Special Interest Group on Pattern Recognition and
       Understandings, Technical Report No.\ PRU89-65,
       1989 (in Japanese).
 \item E. Oomoto, K. Tanaka, H. Nunokawa, T. Kunishima, M. Yoshikawa,
       and Y. Masunaga.
       An Experiment of Collaborative Electronic Publishing --- Design
       and Implementation of Database Workbook ---.
       IPSJ Special Interest Group on Database Systems, Technical Report 
       No.\ DBS99-17,
       July 1994 (in Japanese).
 \item T. Kunishima and Y. Kambayashi.
       Workflow Model on a Workflow Management System WorkFlowBase.
       IPSJ Special Interest Group on Database Systems, Technical Report 
       No.\ DBS104-41,
       July 1995 (in Japanese).
 \item S. Uemura, J. Adachi, M. Arikawa, Y. Kiyoki, S. Shimojo, T. Kato, 
       I. Kojima, K. Saisho, and T. Kunishima.
       Researches on Multimedia Information-base Technology.
       IPSJ Special Interest Group on Database Systems, Technical Report 
       No.\ DBS106-21,
       Jan. 1996 (in Japanese).
 \item K. Yokota, T. Kunishima, and H. Nakanishi.
       Hypothetical Query Processing for Workflow Management.
       Proceedings of the Seminar on Logical Databases and the Meaning
       of Change (UPMAIL Technical Report No.\ 129, Uppsala University),
       pp.\ 44--49, Dagstuhl, Germany, Sep.\ 1996.
\end{enumerate}

\section*{Convention Records}
\begin{enumerate}
 \item T. Kunishima and Y. Kambayashi.
       Applying a Logical Language {HILOG-R} to {CAD} Databases.
       43th National Convention Record of IPSJ, Vol.\ 4, pp.\ 158-159,
       Oct.\ 1991 (in Japanese).
 \item T. Kunishima, T. Furukawa, and Y. Kambayashi.
       Interactive Support for Designing Object-Oriented Databases.
       44th National Convention Record of IPSJ, Vol.\ 4, pp.\ 169-170,
       Mar.\ 1992 (in Japanese).
 \item T. Kunishima and Y. Kambayashi.
       Schedule Management in a Computer-Supported Cooperative Work
       Environment based on Database Systems.
       45th National Convention Record of IPSJ, Vol.\ 6, pp.\ 281-282,
       Oct.\ 1992 (in Japanese).
 \item T. Kunishima, T. Shimada, and Y. Kambayashi.
       Implementation and Applications of Generalized Trigger
       Mechanisms for VirtualOffice.
       46th National Convention Record of IPSJ, Vol.\ 6, pp.\ 247-248,
       Mar.\ 1993 (in Japanese).
 \item T. Kunishima and Y. Kambayashi.
       A Description Method for Cooperative Activities using Dynamic
       Finite Automata.
       46th National Convention Record of IPSJ, Vol.\ 6, pp.\ 249-250,
       Mar.\ 1993 (in Japanese).
 \item T. Kunishima and Y. Kambayashi.
       An Inheritance Mechanism on Aggregation Hierarchies of
       Cooperative Works.
       47th National Convention Record of IPSJ, Vol.\ 6, pp.\ 283-284,
       Oct.\ 1993 (in Japanese).
 \item T. Kunishima and Y. Kambayashi.
       An Execution Method for Workflows Based on Activity Objects.
       48th National Convention Record of IPSJ, Vol.\ 6, pp.\ 241-242,
       Mar.\ 1994 (in Japanese).
 \item T. Kunishima and Y. Kambayashi.
       A View Function on Workflow Model.
       49th National Convention Record of IPSJ, Vol.\ 6, pp.\ 247-248,
       Sept.\ 1994 (in Japanese).
 \item T. Kunishima and Y. Kambayashi.
       Implementation of A Prototype for A WorkFlow Management System
       ``WorkFlowBase''.
       50th National Convention Record of IPSJ, Vol.\ 6, pp.\ 183-184,
       Mar.\ 1995 (in Japanese).
 \item T. Kunishima and Y. Kambayashi.
       A Study for Openness of Workflow Management Systems.
       51th National Convention Record of IPSJ, Vol.\ 6, pp.\ 179-180,
       Sept.\ 1995 (in Japanese).
 \item N. Segawa, M. Hiji, T. Kunishima, and K. Watanabe.
       Description in the Frame of Dynamism of Human Communications.
       51th National Convention Record of IPSJ, Vol.\ 6, pp.\ 187-188,
       Sept.\ 1995 (in Japanese).
\end{enumerate}