%#!latex main

\chapter{A Workflow Model with Database Technologies}
\label{chap:wfbase}

In this chapter a formal workflow data model is proposed.  This model
represents a workflow as a \emph{workflow template} (WFT), a set of
\emph{activity objects}.  Each activity object corresponds to each
office work units of the workflow.  An activity object is regarded as a
transformation function of \emph{message objects}, executed by an agent
or by sub activity objects as subroutines.  Two kinds of flows,
\emph{horizontal flows} and \emph{vertical flows}, are derived from
input-output relationships and part-of relationships between activity
objects, respectively.  An execution model of a WFT is defined as a
production system which treats horizontal flows and vertical flows as
production rules and definite clauses, respectively.

We also propose an instantiation mechanism of workflows in a formal way.
A partial ordering of WFTs is defined based on Smyth orderings of its
components.  Instantiation of a WFT is treated as an assignment over this 
ordering.

Based on these formal models, a workflow database, \emph{workflow base},
is formerly defined.  We also discuss about its integrity constraints.

%\section{Motivations}
%\label{sec:wfbase:intro}

\input{yokota3}

\section{Summary}
\label{sec:wfbase:conclu}

In this chapter a formal workflow data model, \emph{workflow base} was
proposed.  This model represents a workflow as a \emph{workflow
template} (WFT), a set of \emph{activity objects}.  Each activity object
corresponds to each office work units of the workflow, executed by an
agent or by sub activity objects as subroutines.  

Two kinds of flows, \emph{horizontal flows} and \emph{vertical flows},
are derived from input-output relationships and part-of relationships
between activity objects, respectively.  An execution model of a WFT is
defined as a production system which treats horizontal flows and
vertical flows as production rules and definite clauses,
respectively.  This execution model deals with both static flow control
and dynamic dispatching of subworks in the same matter.

We also proposed an instantiation mechanism of workflows in a formal way.
A partial ordering of WFTs is defined based on Smyth orderings of its
components.  Instantiation of a WFT is treated as an assignment over this 
ordering.

Based on these formal models, a workflow database, \emph{workflow base},
was formerly defined.  We also discussed about its integrity constraints.

